Estimating nucleon substructure properties in a\\[.2ex] unified model of p-Pb and Pb-Pb collisions

### 2018-05-10 talk/qm2018.tex
add intro slide to bayesian parameter estimation

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \Large I. Bayesian parameter estimation of heavy-ion collisions
  \end{center}
\end{frame}

@@

    \node<2>[xshift=1cm, yshift=2.1cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<2>[anchor=north east, align=center, inner sep=.6cm] (post_p) at (current page.north east){
        \includegraphics[width=.6\textwidth]{posterior_p}\\[1ex]
        \small Initial entropy deposition\\
        \small consistent with geometric mean\\[1ex]
        $dS/dy \sim \sqrt{T_A T_B}$
    };
    \node<3>[xshift=2.7cm, yshift=0.5cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<3>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
        \includegraphics[width=.5\textwidth]{posterior_parton_number}\\[2ex]
        No strong preference for a specific\\
        constituent number
    };
    \node<3>[align=center] (disfavor) at (6, 2) {
        One constituent\\
        (no substructure) disfavored
    };
    \draw<3>[->, thick, theme] (disfavor.south) to [out=-90, in=90] (7.55, -1.2);
    \node<4>[xshift=2.15cm, yshift=0cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<4>[anchor=east] (proton) at (current page.east) {
      \includegraphics[width=.5\textwidth]{proton_posterior_shape}
    };
    \node<4>[xshift=-3.2cm, yshift=-1.6cm, align=right] (proton_label) at (proton.north west) {
      Informative posterior on constituent\\
      sampling radius $r$ and width $w$\\[2ex]
      $r=0.99_{-0.15}^{+0.17}$~fm\\[1ex]
      $w=0.47_{-0.15}^{+0.20}$~fm
    };
    \node<5>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
      \includegraphics[width=.6\textwidth]{region_shear}
    };
    \node<6>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
      \includegraphics[width=.6\textwidth]{region_bulk}
    };
  \end{tikzpicture}
\end{frame}

%\begin{frame}{\scshape QGP medium properties}
%  \begin{figure}
%    \includegraphics[width=.9\textwidth]{region_shear_bulk}
%  \end{figure}
%  \begin{center}
%    Pb-Pb at 2.76 and 5.02 TeV, better constraint on $\eta/s$ than p-Pb and Pb-Pb at 5.02 TeV
%  \end{center}
%\end{frame}

\begin{frame}[plain]
  \begin{columns}
    \begin{column}{.5\textwidth}
      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
      \begin{tabular}{ll}
        norm           & 20.7 \\
        $p$            & 0.0314 \\
        $\sigma_\text{fluct}$  & 0.835 \\
        $r$      & 0.903 \\
        $n$ partons  & 6.48 \\
        $\chi$  & 0.288 \\
        $d_\text{min}^3$          & 1.11 \\
        $\tau$ fs         & 0.595 \\
        $\eta/s$ min       & 0.122 \\
        $\eta/s$ slope     & 0.742 \\
        $\eta/s$ crv       & -0.203 \\
        $\zeta/s$ max      & 0.0510 \\
        $\zeta/s$ width    & 0.0275 \\
        $\zeta/s$ $T_0$       & 0.185 \\
        $T_\text{sw}$        & 0.143
      \end{tabular}
    \end{column}
    \begin{column}{.4\textwidth}
      \includegraphics[height=.95\textheight]{find_map}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{\scshape Emulator validation}
  \begin{columns}
    \begin{column}{.35\textwidth}
      {\scshape Preprocessing}
      \begin{enumerate}
        \item subtract mean and divide out variance
        \item project onto principle component basis
      \end{enumerate}
      \medskip
      {\scshape k-fold cross validation}
      \begin{itemize}
        \item Subdivide design into $k$ partitions. Choose one for
          validation and use the rest for training. Repeat for all partitions.
      \end{itemize}
    \end{column}
    \begin{column}{.65\textwidth}
      \includegraphics[width=\textwidth]{validation_example}
    \end{column}
  \end{columns}
\end{frame}

@@

>>>>

\begin{frame}{Challenge of rigorous model-to-data comparison}
  \begin{figure}
    \begin{tikzpicture}
        \node [anchor=east, left=1cm  of current page.center, yshift=1.5cm]
          (t1) {\large \color{theme} Parameters};
        \node [anchor=east, left=1cm  of current page.center, yshift=1cm]
            (a1) {\small shear viscosity};
        \node [anchor=east, left=1cm  of current page.center, yshift=.5cm]
            (a2) {\small bulk viscosity};
        \node [anchor=east, left=1cm  of current page.center, yshift=0cm]
            (a3) {\small pre-equilibrium flow};
        \node [anchor=east, left=1cm  of current page.center, yshift=-.5cm]
            (a4) {\small nucleon size};
        \node [anchor=east, left=1cm  of current page.center, yshift=-1cm]
            (a5) {\small nucleon shape};

        % proton-lead observables
        \node [anchor=west, right=1cm of current page.center, yshift=2.5cm]
          (t2) {\large \color{theme} p-Pb observables};
        \node [anchor=west, right=1cm of current page.center, yshift=2cm]
            (b1) {\small charged particle yields};
        \node [anchor=west, right=1cm of current page.center, yshift=1.5cm]
            (b2) {\small mean $p_T$};
        \node [anchor=west, right=1cm of current page.center, yshift=1cm]
            (b3) {\small elliptic flow};
        \node [anchor=west, right=1cm of current page.center, yshift=.5cm]
            (b4) {\small triangular flow};

        % lead-lead observables
        \node [anchor=west, right=1cm of current page.center, yshift=-.5cm]
          (t3) {\large \color{theme} Pb-Pb observables};
        \node [anchor=west, right=1cm of current page.center, yshift=-1cm]
            (c1) {\small charged particle yields};
        \node [anchor=west, right=1cm of current page.center, yshift=-1.5cm]
            (c2) {\small mean $p_T$};
        \node [anchor=west, right=1cm of current page.center, yshift=-2cm]
            (c3) {\small elliptic flow};
        \node [anchor=west, right=1cm of current page.center, yshift=-2.5cm]
            (c4) {\small triangular flow};

        \draw [offblack, line width=0.2pt] (a1.east) -- (b1.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (b2.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (b3.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (b4.west);

        \draw [offblack, line width=0.2pt] (a2.east) -- (b1.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (b2.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (b3.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (b4.west);

        \draw [offblack, line width=0.2pt] (a3.east) -- (b1.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (b2.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (b3.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (b4.west);

        \draw [offblack, line width=0.2pt] (a4.east) -- (b1.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (b2.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (b3.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (b4.west);

        \draw [offblack, line width=0.2pt] (a5.east) -- (b1.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (b2.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (b3.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (b4.west);

        \draw [offblack, line width=0.2pt] (a1.east) -- (c1.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (c2.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (c3.west);
        \draw [offblack, line width=0.2pt] (a1.east) -- (c4.west);

        \draw [offblack, line width=0.2pt] (a2.east) -- (c1.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (c2.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (c3.west);
        \draw [offblack, line width=0.2pt] (a2.east) -- (c4.west);

        \draw [offblack, line width=0.2pt] (a3.east) -- (c1.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (c2.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (c3.west);
        \draw [offblack, line width=0.2pt] (a3.east) -- (c4.west);

        \draw [offblack, line width=0.2pt] (a4.east) -- (c1.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (c2.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (c3.west);
        \draw [offblack, line width=0.2pt] (a4.east) -- (c4.west);

        \draw [offblack, line width=0.2pt] (a5.east) -- (c1.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (c2.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (c3.west);
        \draw [offblack, line width=0.2pt] (a5.east) -- (c4.west);
    \end{tikzpicture}
  \end{figure}
  \begin{center}
    Many correlated input parameters---cannot calibrate by hand!
  \end{center}
\end{frame}

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \Large I. Bayesian parameter estimation
  \end{center}
\end{frame}

@@

    \node<2>[xshift=1cm, yshift=2.1cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<2>[anchor=north east, align=center, inner sep=.6cm] (post_p) at (current page.north east){
        \includegraphics[width=.6\textwidth]{posterior_p}\\[1ex]
        \small Initial entropy deposition\\
        \small consistent with geometric mean\\[1ex]
        $dS/dy \sim \sqrt{T_A T_B}$
    };
    \node<3>[xshift=2.7cm, yshift=0.5cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<3>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
        \includegraphics[width=.5\textwidth]{posterior_parton_number}\\[2ex]
        No strong preference for a specific\\
        constituent number
    };
    \node<3>[align=center] (disfavor) at (6, 2) {
        One constituent\\
        (no substructure) disfavored
    };
    \draw<3>[->, thick, theme] (disfavor.south) to [out=-90, in=90] (7.55, -1.2);
    \node<4>[xshift=2.15cm, yshift=0cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<4>[anchor=east] (proton) at (current page.east) {
      \includegraphics[width=.5\textwidth]{proton_posterior_shape}
    };
    \node<4>[xshift=-3.2cm, yshift=-1.6cm, align=right] (proton_label) at (proton.north west) {
      Informative posterior on constituent\\
      sampling radius $r$ and width $w$\\[2ex]
      $r=0.99_{-0.15}^{+0.17}$~fm\\[1ex]
      $w=0.47_{-0.15}^{+0.20}$~fm
    };
    \node<5>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
      \includegraphics[width=.6\textwidth]{region_shear}
    };
    \node<5>[xshift=5.7cm, yshift=-2.6cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
    \node<6>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
      \includegraphics[width=.6\textwidth]{region_bulk}
    };
    \node<6>[xshift=7.3cm, yshift=-4.2cm] {
      \includegraphics[scale=.065]{magnifying-glass}
    };
  \end{tikzpicture}
\end{frame}

\begin{frame}{\scshape Concluding remarks}
  \vfill
  \begin{enumerate}
    \item Calibrated model simultaneously describes p-Pb and Pb-Pb bulk observables\\[.5ex]
    \item Nucleon substructure necessary to describe p-Pb and Pb-Pb flows (using hydro)\\[.5ex]
    \item p-Pb observables constrain the constituent width and sampling radius, but not the constituent number\\[.5ex]
    \item Varying the beam energy of A-A collisions provides better constraint on temperature dependence of the viscosity than varying system size
  \end{enumerate}
  \begin{flushleft}
    To do: run and analyze events using \textit{maximum a posteriori} parameters\\[2ex]
    \textcolor{theme}{\trento\ with nucleon substructure:}\\
    \url{github.com/morelandjs/trento-partons}\\[1ex]
    \textcolor{theme}{Reproduce my results:}\\
    \url{github.com/morelandjs/hic-param-est-qm18}
  \end{flushleft}
\end{frame}

@@

\begin{frame}{\scshape Emulator validation}
  \begin{columns}
    \begin{column}{.35\textwidth}
      \textcolor{theme}{\scshape k-fold cross validation}\\[.5ex]
      Subdivide design into $k$ partitions. Choose one for validation and use the
      rest for training. Repeat for all partitions.\\[2ex]
      Residuals should sample a unit normal distribution\\
      (far right).
    \end{column}
    \begin{column}{.65\textwidth}
      \includegraphics[width=\textwidth]{validation_example}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}[plain]
  \begin{columns}
    \begin{column}{.5\textwidth}
      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
    \end{column}
    \begin{column}{.4\textwidth}
      \includegraphics[height=.95\textheight]{find_map}
    \end{column}
  \end{columns}
\end{frame}
